Local simple connectedness of boundaries of hyperbolic groups

### 2020-04-02 paper.tex
Proof read first lemma of section 3

@@

Given points $x$ and $y$ in $G \union \boundary G$ we will often denote by
$[x,y]$ a geodesic from $x$ to $y$, thought of as a subset of the Cayley graph
together with its boundary. Geodesics in hyperbolic spaces are not necessarily
unique, so this notation will stand for a choice of geodesic, but the choice
will not matter.

@@

\section{Horoballs in the Rips Complex}\label{sec:horoball}

In Section~\ref{sec:double_dagger} we will make use of a limiting argument,
which will relate the topology of the complement of a point in the boundary
$\boundary G$ to a particular subcomplex of $K$, which we will think of as
a horoball in $K$.  In this section we prove that this subcomplex possesses
certain properties that we will apply later.

@@

For each $n$, let $v_n$ be a vertex of $K_n$ and let $g_n$ be an element of $G$
taken so that $g_n\cdot v_n$ is a point $v \in K_n$ independent of $n$.  Since
$K$ is locally finite, we can pass to a subsequence so that $v_i \in
K_{n_i}$ satisfies the following conditions:
\begin{enumerate}
  \item
    The translates by $g_i$ of large neighbourhoods of $v_i$ in the sphere
    $S_{n_i} \subset G$ are equal: for any $i \geq j$,
    \begin{equation*}
      N_j(v) \intersection g_i(S_{n_i}) = N_j(v) \intersection g_j(S_{n_j})
    \end{equation*}
    (Here $N$ denotes a neighbourhood taken with respect to the word metric on
    $G$.)
  \item
    The sequence $g_i = g_i\cdot 1$ of vertices of $K$ converges to a point
    $\xi \in \boundary G$ as $i\to\infty$.
\end{enumerate}

@@

Let $H$ be the union of the complexes $H_i \subset K$. We view $H$ as a
horosphere, the topology of which approximates that of $\boundary G - \set\xi$.
The rest Section of~\ref{sec:horoball} will be spent justifying this point of view.

\subsection{Geodesics and and the horosphere}

\begin{lemma}\label{lem:geodesicsmeethorosphere}
  Every bi-infinite geodesic with $\xi$ in its limit set meets $H$.
\end{lemma}

\begin{proof}
  Let $\alpha$ be a bi-infinite geodesic with $\xi$ as a limit point. 
  Let the distance $\dist(w,\alpha)$ be realised by a point $x \in \alpha$.

@@

  Since $\dist(g_i, v) = n_i$ for all $i$ and $\dist(x,y_i) \to
  \infty$ as $i\to\infty$, we have the following limit.
  \begin{align*}\label{eqn:zi_closer_to_gi_than_x}
    n_i - \dist(g_i, z_i) & \geq n_i - \dist(g_i, y_i) - \delta \\
                          &= n_i - \dist(g_i, x) + \dist(x, y_i) - \delta\\
                          &\geq \dist(x, y_i) - \dist(v,x) - \delta\\
                          &\to \infty\text{ as $i \to \infty$}\tag{$\star$}
  \end{align*}
  Therefore, for $i$ sufficiently large, there exists a point $p_i$ on $\alpha$
  such that $z_i$ lies between $\xi$ and $p_i$ on $\alpha$ and so that
  $\dist(g_i,p_i)=n_i$. We now show that $\dist(v, p_i)$ is uniformly bounded.

  The limit~\eqref{eqn:zi_closer_to_gi_than_x}, taken together with the fact
  that $\dist(g_i, p_i) = n_i$ and $\dist(g_i, x) \geq n_i - \dist(x, v)$,
  shows that the distances $\dist([z_i, g_i], x)$ and $\dist([z_i, g_i], p_i)$
  tend to infinity as $i\to\infty$.  We now divide into cases according to
  which of $x$ and $p_i$ is closer to $z_i$.
  \begin{enumerate}
    \item If $x$ is between $z_i$ and $p_i$ on $\alpha$ then by considering a
      geodesic triangle with vertices $g_i$, $z_i$ and $p_i$ we see that for
      $i$ sufficiently large, $x$ is within a distance $\delta$ of a point
      $q_i$ on the geodesic from $g_i$ to $p_i$.  Therefore,
      \begin{align*}
        \dist(g_i,x)&\leq \dist(g_i,q_i)+\delta\\
                        &\leq \dist(g_i,p_i)-\dist(p_i,q_i)+\delta\\
                        &\leq \dist(g_i,p_i)-\dist(x,p_i)+2\delta.
      \end{align*}
      It follows that $\dist(x,p_i) \leq \dist(g_i,p_i)-\dist(g_i,x)+2\delta
      = 2\delta$.
    \item If $p_i$ is between $z_i$ and $x$ on $\alpha$ then by an identical 
      argument with the r\^oles of $x$ and $p_i$ reversed we can similarly 
      deduce that 
      \begin{equation*}
        \dist(x,p_i)\leq\dist(g_i,x) - \dist(g_i,p_i) + 2\delta =2\delta.
      \end{equation*}
  \end{enumerate}

  Either way, $\dist(v, p_i) \leq \dist(v, x) + 2\delta$. In summary, we have
  shown that for $i$ sufficiently large there is a point $p_i\in\alpha$ with
  $\dist(g_i,p_i)=n_i$ and $\dist(p_i,f(\Vertices J))$ bounded by $\dist(x,w) +
  2\delta$. Then $p_i \in H_i$ whenever $i$ is at least $\dist(x,w) + 2\delta$.
\end{proof}

>>>>

We call an isometric embedding of an interval into the Cayley graph a
\emph{geodesic}. If the interval is $(-\infty, \infty)$ we call the geodesic a
\emph{bi-infinite geodesic}; if the interval is $[0, \infty)$ we call it a
\emph{geodesic ray.} A bi-infinite geodesic has two ideal end points in
$\boundary G$, while a geodesic ray has one end point in $G$ and one ideal end
point in $\boundary G$.  Given points $x$ and $y$ in $G \union \boundary G$ we
will often denote by $[x,y]$ a geodesic from $x$ to $y$, thought of as a subset
of the Cayley graph together with its boundary. Geodesics in hyperbolic spaces
are not necessarily unique, so this notation will stand for a choice of
geodesic, but the choice will not matter.

@@

@@

\section{Horoballs in the Rips complex}\label{sec:horoball}

In Section~\ref{sec:double_dagger} we will use a limiting argument, which will
relate the topology of the complement of a point in the boundary $\boundary G$
to a particular subcomplex of $K$, which we will think of as a horoball in $K$.
In this section we prove that this subcomplex possesses the properties that we
will apply later.

@@

For each $n$, let $v_n$ be a vertex of $K_n$ and let $g_n$ be an element of $G$
taken so that $g_n\cdot v_n$ is a point $v \in K_n$ independent of $n$.  Since
$K$ is locally finite, we can pass to a subsequence so that $v_i \in
K_{n_i}$ satisfies the following conditions:
\begin{enumerate}
  \item
    The translates by $g_i$ of large neighbourhoods of $v_i$ in the sphere
    $S_{n_i} \subset G$ are equal: for any $i \geq j$,
    \begin{equation*}
      N_j(v) \intersection g_i(S_{n_i}) = N_j(v) \intersection g_j(S_{n_j})
    \end{equation*}
    (Here $N$ denotes a neighbourhood taken with respect to the word metric on
    $G$.)
  \item
    The sequence $(g_i)_{i=1}^\infty$ of group elements, considered also as a
    sequence of vertices of $K$, converges to a point $\xi \in \boundary G$ as
    $i\to\infty$.
\end{enumerate}

@@

Let $H$ be the union of the subcomplexes $H_i \subset K$. We view $H$ as a
horosphere, the topology of which approximates that of $\boundary G - \set\xi$.
The rest Section of~\ref{sec:horoball} will be spent justifying this point of view.

\subsection{Geodesics and the horosphere}

\begin{lemma}\label{lem:geodesicsmeethorosphere}
  Every bi-infinite geodesic with one end point equal to $\xi$ meets $H$.
\end{lemma}

\begin{proof}
  Let $\alpha$ be a bi-infinite geodesic with $\xi$ as an end point.  Let the
  distance $\dist(v,\alpha)$ be realised by a point $x \in \alpha$.

@@

  Since $\dist(g_i, v) = n_i$ for all $i$ and $\dist(x,y_i) \to
  \infty$ as $i\to\infty$, we have the following limit.
  \begin{align*}\label{eqn:zi_closer_to_gi_than_x}
    n_i - \dist(g_i, z_i) & \geq n_i - \dist(g_i, y_i) - \delta \\
                          &= n_i - \dist(g_i, x) + \dist(x, y_i) - \delta\\
                          &\geq n_i - \dist(g_i, v) - \dist(x, v) + \dist(x,y_i) - \delta\\
                          &= \dist(x,y_i) - \dist(x,v) - \delta\\
                          &\to \infty\text{ as $i \to \infty$}\tag{$\star$}
  \end{align*}
  In particular, $n_i > \dist(g_i, z_i)$ for $i$ sufficiently large, so for
  $i$ sufficiently large, there exists a point $p_i$ on $\alpha$
  such that $z_i$ lies between $\xi$ and $p_i$ on $\alpha$ and so that
  $\dist(g_i,p_i)=n_i$. We now show that $\dist(v, p_i)$ is uniformly bounded.

  The limit~\eqref{eqn:zi_closer_to_gi_than_x}, taken together with the fact
  that $\dist(g_i, p_i) = n_i$ and $\dist(g_i, x) \geq n_i - \dist(x, v)$,
  shows that the distances $\dist([z_i, g_i], x)$ and $\dist([z_i, g_i], p_i)$
  tend to infinity as $i\to\infty$.  We now divide into cases according to
  which of $x$ and $p_i$ is closer to $z_i$.
  \begin{enumerate}
    \item If $x$ is between $z_i$ and $p_i$ on $\alpha$ then by considering a
      geodesic triangle with vertices $g_i$, $z_i$ and $p_i$ we see that for
      $i$ sufficiently large, $x$ is within a distance $\delta$ of a point
      $q_i$ on the geodesic from $g_i$ to $p_i$.  Therefore,
      \begin{align*}
        \dist(g_i,x)&\leq \dist(g_i,q_i)+\delta\\
                        &\leq \dist(g_i,p_i)-\dist(p_i,q_i)+\delta\\
                        &\leq \dist(g_i,p_i)-\dist(x,p_i)+2\delta.
      \end{align*}
      It follows that $\dist(x,p_i) \leq \dist(g_i,p_i)-\dist(g_i,x)+2\delta
      \leq \dist(x, v) + 2\delta$.
    \item If $p_i$ is between $z_i$ and $x$ on $\alpha$ then by an identical 
      argument with the r\^oles of $x$ and $p_i$ reversed we can similarly 
      deduce that 
      \begin{equation*}
        \dist(x,p_i)\leq\dist(g_i,x) - \dist(g_i,p_i) + 2\delta \leq  \dist(x,v) + 2\delta.
      \end{equation*}
  \end{enumerate}

  Either way, $\dist(v, p_i) \leq 2\dist(x, v) + 2\delta$. In summary, we have
  shown that for $i$ sufficiently large there is a point $p_i\in\alpha$ with
  $\dist(g_i,p_i)=n_i$ and $\dist(v, p_i)$ bounded by $2\dist(x,w) +
  2\delta$. Then $p_i \in H_i \intersection \alpha$ whenever $i$ is at least
  $2\dist(x,w) + 2\delta$. 
\end{proof}
